In recent decades, with the rise of small portable devices for applications such as the Internet of Things \cite{5579543}
and CubeSats \cite{heidt2000cubesat}, as well as the concepts of Edge Computing \cite{7488250}, Industry 4.0 \cite{lasi2014industry}, 
and Big Data, part of the information must be processed on the \textcolor{blue}{client} device, alleviating the network's load.

To achieve this goal, processors must be more efficient in executing specific operations and attending tasks with lower latency. 
Since most portable devices have a general-purpose processor, it is not possible to run specific applications efficiently. 
Therefore, some solutions must be found to improve its performance. One option is the customization of the processor, 
identifying the algorithms that are most used in each case. For example, cryptography and 
\textcolor{blue}{error-correction codes} are present in almost any portable device today, and many of these algorithms are based on finite field arithmetic.  
Because of this, the acceleration of finite field arithmetic $GF(2^m)$ proceeds to have a significant role \cite{8192501}. 
It is generally used to encode and decode in a communication channel and detect errors in the transmitted data 
(BCH, Reed-Solomon codes \cite{wicker1999reed}). It is also used in asymmetric cryptography, such as Elliptical Curve Cryptography \cite{washington2008elliptic}, 
which is extensively used in the authentication process to exchange private keys or symmetric cryptography 
inside a secure communication channel as Advanced Encryption Standard (AES) \cite{heron2009advanced}.


Furthermore, post-quantum cryptography algorithms (PQC) \cite{8791343} have been developed in recent years 
with the rise of quantum computing research. Some of the survivors of the third round of the NIST's PQC competition \cite{moody2016post} 
use $GF(2^m)$ arithmetic (Classic McEliece \cite{bernstein2017classic}, Rainbow \cite{10.1007/11496137_12}, HQC \cite{melchor2018hamming}, and GeMSS \cite{casanova2017gemss}). 
These algorithms require high computing power to generate the keys and encrypt/decrypt the data, 
becoming the main bottleneck in processing for small devices, such as IoT end-nodes \cite{McGinthy2018} and Cubesats \cite{Maheshwarappa2017}.


Traditionally, there are three ways to optimize a processor. The first solution is adding coprocessors 
to the system to perform those specific operations. The second is expanding the base instructions 
of the computer architecture. Moreover, the third solution is to make a hybrid system \cite{4352011} between the first and 
the second solution, adding coprocessors and specific instructions depending on the case. 
We focus on the second solution, expanding the RISC-V base instruction set (RV32I) \textcolor{blue}{due to its simplicity to integrate within a core}. 


This work aims to propose a flexible instruction set for RV32 cores
capable of accelerating any algorithm based on finite field arithmetic $GF(2^m)$, thus improving processor performance. A tradeoff 
between the base instruction set and specific custom instructions can be found for applications that require high flexibility and, 
at the same time, acceleration in the CPU execution time. For example, a portable device generally contains different error-correction codes 
and cryptographic standards or proprietary ciphers.


The rest of the paper is organized as follows. Section 2 describes the mathematical basis of $GF(2^m)$, previous related works, 
and contributions of this work. Section 3 introduces the instruction set extension. Section 4 shows the simulations and 
experimental results in a RISC-V SoC (SweRVolf). Finally, section 5 finishes with the conclusion of this work.


